Accesible, realistic genome simulation with selection using \stdpopsim

### 2025-02-27 analysis2_manuscript.tex

    % simulation
    To meet the growing need for the interpretation,
    analysis, and exploration of realistic and complex evolutionary models
    the field of population genetics has increasingly turned to simulation.
    Simulation in population genetics has a long history 
    including both backward-in-time coalescent simulations
    \citep{kingman1982genealogy,hudson1983testing, hudson1990gene}
    and forward-in-time simulations of complex demography and selection
    \citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.  \kelcomment{maybe also cite https://doi.org/10.1093/bioinformatics/btn522}
    The development of simulation tools has been driven by the need to
    understand the effects of complex evolutionary processes on genetic
    variation \citep[e.g.,][]{galloway2020few},\kelcomment{also cite: PMID: 32487519 } to provide a null model for hypothesis testing
    \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
    to explore the power and limitations of statistical methods \citep[e.g.,][]{przeworski2002signature},
    and increasingly to provide a basis for machine learning and other
    simulation-based inference methods \citep[e.g.,][]{beaumont2002approximate,pavlidis2010searching,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic}. \kelcomment{also cite: PMID: 36617238 }
    While this is so, joint simulation of complex demography and selection
    is challenging,  requires a deep understanding of the underlying
    evolutionary processes, and necessitates a number of parameter choices including
    the strength of selection and the
    recombination rate, as well as a parameterized demographic model.
    This is a daunting task for many researchers, and can be a barrier to
    the adoption of simulation-based methods in population genetics.
    Furthermore, different simulation and inference tools may use different structures
    and parameter scalings for their models.

    % stdpopsim
    In light of the complexities associated with estimating and simulating
    population genetic models, it is no surprise that
    a lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers \citep[e.g.,][]{ragsdale2020lessons}.
    This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
    While that is so, the original version of \stdpopsim did not include
    models of selection, which is a major limitation for more empirical
    applications of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding processes such
    as adaptation, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
    background selection on genetic diversity
    \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.
    Ideally, one would like
    to have a single, unified framework for simulating both neutral and
    non-neutral evolutionary processes, and to be able to compare the
    results of these simulations to empirical data in a manner that is
    both accessible to a wide range of researchers and highly reproducible. 
    Further the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the size and density of functional elements, and
    non-equilibrium demographic histories. Ideally, one should be able
    to model all of these features using estimates/annotations obtained
    from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.

@@

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for curated addition of new
    species and models, and can use either
    msprime \citep{Baumdicker2022} or SLiM \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    \igcomment{edited here to add details on the spec of the DFE and Annotation classes
    moved below the information on how they are used in simulation. Please check!}
    To extend \stdpopsim's functionality to include models of selection, 
    we introduce two new classes
    %an important new class representing a specfic distribution of fitness effects
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where 
    each mutation type is associated with a dominance coefficient (in $[0,1]$),
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    Each \texttt{Annotation} object is typically associated with a collection of functional
    genomic elements specified using a GFF3 file. %
    %    
    For example, say that we wish to implement a model of selection consistent with the
    DFE inferred by \citep{kim2017inference} for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonimous sites.
    Their proportions are set to $0.3$ and $0.7$ respectively, following their expected
    prevelanece in exons.
    The neutral mutation type is associted with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by \citep{kim2017inference}.
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients
    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below).

@@

    \igcomment{modified this paragraph to focus on the catalog and moved simulation details to next paragraph}
    As with other \stdpopsim models, when incorporating selection we aim to hew as closely as possible
    to realistic genomes.
    Thus, the \stdpopsim catalog provides \texttt{Annotation} object based on public annotations
    of specices' functional genomic elements 
    and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
    Since the vast majority of published DFE estimates are based on a limited number of species,
    \stdpopsim currently has implemented DFEs for four species:
    \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
    humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While the DFEs are from a limited number of sepecies, these DFEs can be applied to other species
    in the catalog. For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Furthermore, the user can specify a custom DFE and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have 
    a published DFE included in the catalog. This flexibility allows for a wide range of
    models of selection to be simulated. %, and for the results of these simulations to be
    %compared to empirical data or used as a null model for hypothesis testing.
    % DRS: commenting that stuff out because there are obviously other use cases that could be
    % listed here (e.g. ML/ABC, and if we don't know enough about a species to do that then
    % we probably have no business doing hypothesis testing either, which could be just as
    % misleading or maybe even more so), so the simplest solution is to cut this as it is
    % not needed here.

    From a user's perspective, a model of selection is specified by pairing a \texttt{DFE} object
    with a collection of genomic segments defined using an \texttt{Annotation} object.
    Multiple such pairings can be defined in one simulation
    to provide a rich model of how selection may vary along a chromosome.
    For example, coding sequences may be associated with one DFE and non-coding sequences in exons with another DFE.
    Simulations with selection are then implemented by specifying a species, 
    a portion of the genome to simulate,
    a genetic map if available, a model of demography, and pairings of DFEs with genomic segments
    using either the python API or the command line interface (Figure \ref{fig:schematic}B-C).
    %IG: commented this out because I think it's covered by the text I wrote above
    %A \texttt{DFE} object can apply to specified intervals within a chromosome,
    %and can be combined with other \texttt{DFE} objects 
    Simulations with selection can currently be generated only 
    using the SLiM simulation engine, which is
    %is perhaps
    the most flexible simulation engine for modeling selection available to-date
    (msprime currently cannot be used to simulate models of selection).
    %IG: this is a nice statement, but I couldn't figure where to plug it after my changes so I'm leaving it commented
    %all in an automated and user-friendly fashion.

@@

    \section*{Inference of $N_e(t)$ in the context of selection}
    One of the most common applications of population genetic inference is to estimate
    the effective population size over time, $N_e(t)$, from genetic data. This can be done
    using a variety of methods, including the sequentially Markovian coalescent
    \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
    \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
    through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
    as well as through identity by descent information \citep{santiago2020recent}.
    Since all of these methods assume neutral evolution, they are typically applied to genomic
    segments that are not expected to be directly affected by selection (e.g., by masking exons
    and conserved elements).
    However, natural selection acting on linked sites (that are masked out of the analysis)
    can bias estimates of $N_e(t)$
    away from true census population sizes by increasing the 
    rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}. 
    Using \stdpopsim, we can easily examine the effect of selection on linked sites,
    which is commonly referred to as \emph{background selection}, 
    on different methods for inferring $N_e(t)$.

@@

    \begin{figure}[b!]
        \centering
        %IG: NEED TO LINK NEW FIG HERE
        %\includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
        \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}
        \caption{
        \label{fig:human-demography}
        Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
        with and without purifying selection on exons. 
        Rows correspond to the three extant populations in the simulation: CHB, CEU, and YRI.
        Columns correspond to the four inference methods:    
        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
        Each plot depicts the inferred $N_e(t)$ on the three datasets simulated without selection (blue)
        and the three datasets simulated with the background effects of purifying selection on exons (orange),
        alongside the true values of $N_e$ used in simulation (black).
        }
    \end{figure}

@@

    To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
    We simulated 100 genomes under a two-epoch demographic model inferred for vaquita in \textcite{robinson2022critically},
    and a constant recombination rate across the genome (see \textbf{Methods}).
    In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
    annotated from the vaquita genome assembly.
    This DFE model implements a relationship between selection and dominance 
    in which very deleterious mutations (with selection coefficient $s<-0.1$) are fully recessive
    (dominance coefficient $h=0$) and weakly deleterious mutations ($s\geq -0.001$) are nearly
    additive ($h=0.4$; see \textbf{Methods}).
    Here, we see that the site-frequency-based method (\stairway) produces more accurate
    inference than the two methods based on the sequenctioal Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
    Reagrdless of these difference, we see that background selection does not
    considerably influence the inference of any of the four methods also for this species,
    which has a very different evolutionary history than humans.

@@

\section*{Estimation of the Distribution of Fitness effects}
    \label{dfe}
    Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE) of new mutations
    from genetic data. The new framework of modeling selection in \stdpopsim is ideally designed to benchmark and compare
    the performance of different methods for inferring the DFE. 
    %IG: I don't think we need this intro.
    %We performed simulations using the 
    %human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
    %
    A challenge when comparing different inferred and simulated distributions of the selection coefficient ($s$) is differing convention regarding its definition.
    The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$.
     %https://academic.oup.com/genetics/article/207/3/1103/5930676
    Moreover, inferred distributions of selection coefficients are typically scaled by an inferred ancestral
    effective population size $N_e$, and different methods assume different scaling.
    \dadicli infers the distribution of $2 N_e s$ while \polydfe and \grapes infer the distribution of $4 N_e s$.
    %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
    For comparison, here we have normalized all inference results to the convention of SLiM.
    \igcomment{We should add a detailed description in the Methods of what we do to scale the output of each DFE-inference method.}

    % Add citation to fastDFE as the successor to polyDFE: https://academic.oup.com/mbe/article/41/5/msae070/7641109 ?

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
        \igcomment{NEED TO UPDATE FIG WITH NEW VERSION}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons (see \textbf{Methods}).
        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulated datasets,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines.
        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets;
        white bars represent the distribution used in the simulation.
        }
        \label{fig:homsap-dfe.ooa}
    \end{figure}

@@

    %IG: describe experiment and cite figure later
    %In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
    Next, we compared these three DFE-inference methods on simulations of the vaquita porpoise genome.
    Recall that these simulations apply a gamma-dsitributed selection coefficient to mutations in exons
    and a relationship between the selection and dominance coefficients (see \textbf{Methods}).
    In this setting, all methods perform uniformly worse than in the human genome simulations,
    with consistent underestimation of the mean selection coefficient    
    and overestimation of the shape parameter (Figure \ref{fig:vaquita-dfe}A-B). 
    Similar patterns are observed when re-running the simulations with a simpler demographic model
    without population size changes (Figure \ref{fig:vaquita-dfe.constant}).
    We suspect that this is due to the fact that all DFE inference methods used here assume an additive model ($h = \frac{1}{2}$),
    while mutations in the simulated genomes were more recessive ($h < \frac{1}{2}$), and their dominance decreased with the strength of selection. 
    To examine the influence of this model misspecification, we compared the inferred distributions of $s$ to the 
    simulated distribution of $2hs$, which would have been equal to $s$ under an additive model (Figure \ref{fig:vaquita-dfe}C).
    Reassuringly, all three methods infer distributions that appear to fit the main mode of the simulated distribution of $2hs$.
    This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
   \kelcomment{I moved the last sentence from here to the Discussion and fleshed it out}

    %add PhoSin DFE figure here
    \begin{figure}[t!]
        \centering
        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
        \caption{
        \label{fig:vaquita-dfe}
        Inferred versus simulated DFEs from simulated vaquita genomes.
        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$). 
        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; panel A) and shape parameter (panel B).
        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
        \grapes, \polydfe, and \dadicli.
        Panel C shows the simulated distribution of $2 h s$,
        along with the simulated distribution of $s$ and the median inferred DFEs. \igcomment{What is median inferred DFE?}
        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
        }
    \end{figure}

@@

    We compared the overall performance of the three sweep-detection methods using
    receiver operating characteristic (ROC) curves (Figure \ref{fig:roc-curves}).
    Different methods perform better for each of the sampled populations,
    with \sweepfinder performing betteer than the other two on the CEU sample,
    and \diploshic performing better on the YRI sample.
    This is likely a results of the fact that we trained \diploshic on simulations under a constant size population,
    which more closely matches the demographic history of YRI, than that of CEU, which 
    is characterized by a strong recent bottleneck followed by population expansion.
    Interestingly, the simple method that uses reduced diversity ($\pi$) as an indicator
    for a selective sweep performs reasonably well overall (especially on the CEU sample).
    However, its accuracy is much more variable along the chromosome compared to the other two method (see Figure \ref{fig:chr1_power}),
    and for quite a few locations along the chromosome it has power below $50\%$
    when applied to both sample sets.
    In that sense, \sweepfinder appears to have the least variation in power along the chromosome or across populations.

>>>>

    % simulation
    To meet the growing need for the interpretation,
    analysis, and exploration of realistic and complex evolutionary models
    the field of population genetics has increasingly turned to simulation.
    Simulation in population genetics has a long history 
    including both backward-in-time coalescent simulations
    \citep{kingman1982genealogy,hudson1983testing, hudson1990gene}
    and forward-in-time simulations of complex demography and selection
    \citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.  \kelcomment{maybe also cite https://doi.org/10.1093/bioinformatics/btn522}
    The development of simulation tools has been driven by the need to
    understand the effects of complex evolutionary processes on genetic
    variation \citep[e.g.,][]{galloway2020few},\kelcomment{also cite: PMID: 32487519 } to provide a null model for hypothesis testing
    \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
    to explore the power and limitations of statistical methods \citep[e.g.,][]{przeworski2002signature},
    and increasingly to provide a basis for machine learning and other
    simulation-based inference methods \citep[e.g.,][]{beaumont2002approximate,pavlidis2010searching,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic}. \kelcomment{also cite: PMID: 36617238 }
    While this is so, joint simulation of complex demography and selection
    is challenging,  requires a deep understanding of the underlying
    evolutionary processes, and necessitates a number of parameter choices including
    the strength of selection and the
    recombination rate, as well as a parameterized demographic model.
    This is a daunting task for many researchers, and can be a barrier to
    the adoption of simulation-based methods in population genetics.
    Furthermore, different simulation and inference tools may use different structures
    and parameter scalings for their models,
    making it difficult to directly compare across these tools and the studies that use them.

    % stdpopsim
    In light of the complexities associated with estimating and simulating
    population genetic models, it is no surprise that
    a lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers \citep[e.g.,][]{ragsdale2020lessons}.
    This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
    While that is so, the original version of \stdpopsim did not include
    models of selection, which is a major limitation for more empirical
    applications of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding processes such
    as adaptation, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
    background selection on genetic diversity
    \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.
    Ideally, one would like
    to have a single, unified framework for simulating both neutral and
    non-neutral evolutionary processes, and to be able to compare the
    results of these simulations to empirical data in a manner that is
    both accessible to a wide range of researchers and highly reproducible. 
    Further, the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the size and density of functional elements, and
    non-equilibrium demographic histories, and where possible, modeling these features using estimates/annotations obtained
    from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.

@@

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for curated addition of new
    species and models, and can use either
    msprime \citep{Baumdicker2022} or SLiM \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    \igcomment{edited here to add details on the spec of the DFE and Annotation classes
    moved below the information on how they are used in simulation. Please check!}
    To extend \stdpopsim's functionality to include models of selection, 
    we introduce two new classes
    %an important new class representing a specfic distribution of fitness effects
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where 
    each mutation type is associated with a dominance coefficient (in $[0,1]$),
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    Each \texttt{Annotation} object is typically associated with a collection of functional
    genomic elements specified using a GFF3 file. %
    %    
    For example, say we wanted to implement a model of selection consistent with the
    DFE inferred by \citep{kim2017inference} for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonimous sites,
    with their proportions set to following their expected
    prevalence in exons.
    The neutral mutation type is associted with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by \citep{kim2017inference}.
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients
    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below).

@@

    \igcomment{modified this paragraph to focus on the catalog and moved simulation details to next paragraph}
    As with other aspects of \stdpopsim, we intend the selection models to be as biologicaly accurate as possible.
    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on specices' publically available functional genomic elements 
    and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
    There are relatively few species with published DFE estimates in the literature.
    \stdpopsim currently has implemented DFEs for four species in the catalogue:
    \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
    humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While the DFEs are from a limited number of sepecies, these DFEs can be applied to other species
    in the catalog if and when it is suitable. 
    For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Furthermore, the user can specify a custom DFE and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have 
    a published DFE included in the catalog. This flexibility allows for a wide range of
    models of selection to be simulated. %, and for the results of these simulations to be
    %compared to empirical data or used as a null model for hypothesis testing.
    % DRS: commenting that stuff out because there are obviously other use cases that could be
    % listed here (e.g. ML/ABC, and if we don't know enough about a species to do that then
    % we probably have no business doing hypothesis testing either, which could be just as
    % misleading or maybe even more so), so the simplest solution is to cut this as it is
    % not needed here.

    From a user's perspective, a model of selection is specified by pairing a \texttt{DFE} object
    with a collection of genomic segments defined using an \texttt{Annotation} object.
    Multiple such pairings can be defined in one simulation
    to provide a rich model of how selection may vary along a chromosome.
    For example, coding sequences may be associated with one DFE and non-coding sequences in exons with another DFE.
    Simulations with selection are then implemented by specifying a species, 
    a portion of the genome to simulate,
    a genetic map if available, a model of demography, and pairings of DFEs with genomic segments
    using either the python API or the command line interface (Figure \ref{fig:schematic}B-C).
    %IG: commented this out because I think it's covered by the text I wrote above
    %A \texttt{DFE} object can apply to specified intervals within a chromosome,
    %and can be combined with other \texttt{DFE} objects 
    Simulations with selection are generated 
    using the SLiM simulation engine, which is
    %is perhaps
    the most flexible simulation engine for modeling selection available to-date.
    %IG: this is a nice statement, but I couldn't figure where to plug it after my changes so I'm leaving it commented
    %all in an automated and user-friendly fashion.

@@

    \section*{Inference of $N_e(t)$ in the context of selection}
    One of the most common applications of population genetic inference is to estimate
    the effective population size over time, $N_e(t)$, from genetic data. This can be done
    using a variety of methods, including the sequentially Markovian coalescent
    \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
    \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
    through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
    as well as through identity by descent information \citep{santiago2020recent}.
    Since all of these methods assume neutral evolution, they are typically applied to genomic
    segments that are not expected to be directly affected by selection (e.g., by masking exons
    and conserved elements).
    However, natural selection acting on linked sites (even when masked out of the analysis)
    can bias estimates of $N_e(t)$
    away from true census population sizes by increasing the 
    rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}. 
    Using \stdpopsim, we can easily examine the effect of selection on linked sites,
    which is commonly referred to as \emph{background selection}, 
    on different methods for inferring $N_e(t)$.

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
        \caption{
        \label{fig:human-demography}
        %citations aren't always allowed in captions. We should check with MBE.
        Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
        with and without purifying selection on exons. 
        Rows correspond to the three extant populations in the simulation: CHB, CEU, and YRI.
        Columns correspond to the four inference methods:    
        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
        Each plot depicts the inferred $N_e(t)$ on the three datasets simulated without selection (blue)
        and the three datasets simulated with the background effects of purifying selection on exons (orange),
        alongside the true values of $N_e$ used in simulation (black).
        }
    \end{figure}

@@

    To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
    We simulated 100 genomes under a two-epoch demographic model inferred for vaquita in \textcite{robinson2022critically},
    and a constant recombination rate across the genome (see \textbf{Methods}).
    In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
    annotated from the vaquita genome assembly.
    This DFE model implements a relationship between selection and dominance 
    in which very deleterious mutations (with selection coefficient $s<-0.1$) are fully recessive
    (dominance coefficient $h=0$) and weakly deleterious mutations ($s\geq -0.001$) are nearly
    additive ($h=0.4$; see \textbf{Methods}).
    Here, we see that the site-frequency-based method (\stairway) produces more accurate
    inference than the two methods based on the sequenctioal Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
    Reagrdless of these difference, we again see that background selection does not
    considerably influence the inference of any of the four methods also for this species,
    which has a very different evolutionary history than humans.

@@

\section*{Estimation of the Distribution of Fitness effects}
    \label{dfe}
    Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE) of new mutations
    from genetic data. The new framework of modeling selection in \stdpopsim is ideally designed to benchmark and compare
    the performance of different methods for inferring the DFE. 
    %IG: I don't think we need this intro.
    %We performed simulations using the 
    %human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
    %
    A challenge when comparing different inferred and simulated distributions of the selection coefficient ($s$) is differing convention regarding its definition.
    The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$.
     %https://academic.oup.com/genetics/article/207/3/1103/5930676
    Moreover, inferred distributions of selection coefficients are typically scaled by an inferred ancestral
    effective population size $N_e$, and different methods assume different scaling.
    \dadicli infers the distribution of $2 N_e s$ while \polydfe and \grapes infer the distribution of $4 N_e s$.
    %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
    For comparison, here we have normalized all inference results to the convention of SLiM (see \textbf{Methods}).
    % Add citation to fastDFE as the successor to polyDFE: https://academic.oup.com/mbe/article/41/5/msae070/7641109 ?

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons (see \textbf{Methods}).
        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulated datasets,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines.
        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets;
        white bars represent the distribution used in the simulation.
        }
        \label{fig:homsap-dfe.ooa}
    \end{figure}

@@

    %IG: describe experiment and cite figure later
    %In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
    Next, we compared these three DFE-inference methods on simulations of the vaquita porpoise genome.
    Recall that these simulations apply a gamma-distributed selection coefficient to mutations in exons
    and a relationship between the selection and dominance coefficients (see \textbf{Methods}).
    In this setting, all methods perform uniformly worse than in the human genome simulations,
    with consistent underestimation of the mean selection coefficient    
    and overestimation of the shape parameter (Figure \ref{fig:vaquita-dfe}A-B). 
    Similar patterns are observed when re-running the simulations with a simpler demographic model
    without population size changes (Figure \ref{fig:vaquita-dfe.constant}).
    We suspect that this is due to the fact that all DFE inference methods used here assume an additive model ($h = \frac{1}{2}$),
    while mutations in the simulated genomes were more recessive ($h < \frac{1}{2}$), and their dominance decreased with the strength of selection. 
    To examine the influence of this model misspecification, we compared the inferred distributions of $s$ to the 
    simulated distribution of $2hs$, which would have been equal to $s$ under an additive model (Figure \ref{fig:vaquita-dfe}C).
    Reassuringly, all three methods infer distributions that appear to fit the main mode of the simulated distribution of $2hs$.
    This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
   \kelcomment{I moved the last sentence from here to the Discussion and fleshed it out}

    %add PhoSin DFE figure here
    \begin{figure}[t!]
        \centering
        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
        \caption{
        \label{fig:vaquita-dfe}
        Inferred versus simulated DFEs from simulated vaquita genomes.
        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$). 
        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; panel A) and shape parameter (panel B).
        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
        \grapes, \polydfe, and \dadicli.
        Panel C shows the simulated distribution of $2 h s$ (gamma parameters averaged across replicates),
        along with the simulated distribution of $s$.
        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
        }
    \end{figure}

@@

    We compared the overall performance of the three sweep-detection methods using
    receiver operating characteristic (ROC) curves (Figure \ref{fig:roc-curves}).
    Different methods perform better for each of the sampled populations,
    with \sweepfinder performing better than the other two on the CEU sample,
    and \diploshic performing better on the YRI sample.
    This is likely a results of the fact that we trained \diploshic on simulations under a constant size population,
    which more closely matches the demographic history of YRI, than that of CEU, which 
    is characterized by a strong recent bottleneck followed by population expansion.
    Interestingly, the simple method that uses reduced diversity ($\pi$) as an indicator
    for a selective sweep performs reasonably well overall (especially on the CEU sample).
    However, its accuracy is much more variable along the chromosome compared to the other two method (see Figure \ref{fig:chr1_power}),
    and for quite a few locations along the chromosome it has power below $50\%$
    when applied to both sample sets.
    In that sense, \sweepfinder appears to have the least variation in power along the chromosome or across populations.
